---
title: "Reproducible research1"
author: "Michaela Jírù"
date: "14 prosince 2016"
output: html_document
---


#Loading the data
At first, we load libraries, the working directory is set and the data is loaded into R.

```{r message=FALSE}
library(dplyr)
library(xtable)
library(ggplot2)
library(gridExtra)
library(timeDate)
 options(scipen=1, digits=2)
setwd("C:/Users/mjiru/Documents/Coursera/Reproducible research/")
data <- read.csv("activity.csv")
```

#What is mean total number of steps taken per day?

Calculate the total number of steps taken per day 

```{r}
steps_total <- data %>%
    group_by(date) %>%
    summarise(total_steps =sum (steps,na.rm=TRUE))

```
Display few of the results
```{r showtable, results="asis"}
xt <- xtable(head(steps_total))
print(xt, type="html")
```
Display the histogram of steps taken each day
```{r fig.height=5, message=FALSE}
h1 <- ggplot(data= steps_total, aes(total_steps))+geom_histogram(color="black", fill="darkgreen")+
  ylab("Number of occurences")+
  xlab("Total steps taken on a day")+
  ggtitle("Histogram of total steps taken \n on a day.")+
  theme_bw()
h1
```

Calculate the mean and median of the total number of steps taken per day.
```{r}
steps_mean <- steps_total %>%
    summarise(mean_steps =mean (total_steps,na.rm=TRUE),median_steps =median (total_steps,na.rm=TRUE) )
mean <- steps_mean$mean_steps
median <- steps_mean$median_steps

```
The mean value of total number of steps taken per day is **`r mean`** and the median value is **`r median`**.

#What is the average daily activity pattern?
Calculate the average number of steps taken in the 5-minute interval.
```{r}
interval_avg <- data %>%
    group_by(interval) %>%
    summarise(mean = mean(steps, na.rm=TRUE) )

```
Display the results

```{r fig.height=5, message=FALSE}
 a1 <- ggplot(data= interval_avg, aes( interval, mean))+geom_line(color="darkgreen", size=1.2)+
  ylab("Average steps taken")+
  xlab("Interval")+
  ggtitle("Average steps taken during \n all intervals.")+
  theme_bw()
a1
```

Calculate the interval containing the maximum number of steps
```{r}
max <- interval_avg [interval_avg$mean==max(interval_avg$mean),]
max_int <- max$interval
max_value <- max$mean
```

The interval with the highest average number of steps is **`r max_int`**, with the averge number of steps of **`r max_value`**

#Imputing missing values

Calculate the total number of missing values
```{r}
cnt_missing <- nrow(data[is.na(data$steps),])

```

The total number of missing values is **`r cnt_missing`**

Impute the missing values: Use the interval median and create a new dataset.
```{r}
data_imputed <- merge(data, interval_avg, by.x="interval", by.y="interval")
data_imputed$steps <- ifelse( is.na(data_imputed$steps),data_imputed$mean,data_imputed$steps   )


```

Calculate the total number of steps taken each day from new dataset
```{r}
steps_total_imp <- data_imputed %>%
    group_by(date) %>%
    summarise(total_steps =sum (steps))

```

Display the histogram of steps taken each day
```{r fig.height=5, message=FALSE}
h2 <- ggplot(data= steps_total_imp, aes(total_steps))+geom_histogram(color="black", fill="darkgreen")+
  ylab("Number of occurences")+
  xlab("Total steps taken on a day")+
  ggtitle("Histogram of total steps taken \n on a day with imputed values.")+
  theme_bw()

grid.arrange(h1, h2, nrow=1, ncol=2)
```

Calculate the mean and median of the total number of steps taken per day with imputed values.
```{r}
steps_mean_imputed <- steps_total_imp %>%
    summarise(mean_steps =mean (total_steps),median_steps =median (total_steps) )
mean_imp <- steps_mean_imputed$mean_steps
median_imp <- steps_mean_imputed$median_steps

```

The mean value of total number of steps taken per day is **`r mean_imp`** and the median value is **`r median_imp`**.
The difference in mean of the original and imputed data is **`r mean_imp-mean`** and the difference in median is **`r median_imp-median`**

#Are there differences in activity patterns between weekdays and weekends?

Create a factor variable indicating if the day is weekend or weekday.
```{r}
data_imputed$weekend <- isWeekday(data_imputed$date)
data_imputed$weekend<- as.factor(ifelse(data_imputed$weekend, "weekend", "weekday"))

```

Calculate the the average number of steps taken in interval averaged across weekdays or weekends.
```{r}
avg_week <- data_imputed %>%
    group_by(weekend, interval) %>%
    summarise(average =mean (steps))

```

Save weekends and weekdays separately

```{r}
avg_weekdays <-avg_week[avg_week$weekend=="weekday",]
avg_weekend <-avg_week[avg_week$weekend=="weekend",]

```

Display the histogram of steps taken each day
```{r fig.height=5, message=FALSE}
aw1   <- ggplot(data= avg_weekend, aes( interval, average))+geom_line(color="darkgreen", size=1.2)+
  ylab("Average steps taken")+
  xlab("Interval")+
  ggtitle("Average steps taken during \n all intervals at weekends.")+
  theme_bw()

aw2   <- ggplot(data= avg_weekdays, aes( interval, average))+geom_line(color="darkblue", size=1.2)+
  ylab("Average steps taken")+
  xlab("Interval")+
  ggtitle("Average steps taken during \n all intervals at weekdays.")+
  theme_bw()

grid.arrange(aw1, aw2, nrow=1, ncol=2)
```